---
title: "Week Seven: Modeling"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
library(tidyverse)
```

## Activity 6 {.smaller}

```{r eval = F}
http://bit.ly/2oOdqlV # voter_sample.csv
http://bit.ly/2I9CFaC # voter_tidy.csv
```

Data Integrity checks:

1. How have the number of people registered changed over time?
2. What is the distribution of ages within the full file? Within OMV?
3. Did anyone register before age 18?
4. Anyone double registered?

Data Exploration:

1. What is the relationship between party identification and voter proportion?
2. Do metro areas have different party breakdown than rural areas?
3. What is the relationship between age and voter participation when controlling for
party?
4. How do those registered by OMV compare to those not registered by OMV?


##

```{r eval = FALSE}
voter_sample <- read_csv("http://bit.ly/2oOdqlV")
voter_tidy <- read_csv("http://bit.ly/2I9CFaC")
library(lubridate)
voter_tidy %>%
  mutate(birth_year = year(mdy(BIRTH_DATE)),
         age = 2016 - birth_year) %>%
  filter(PARTY_CODE %in% c("REP", "DEM", "NAV")) %>%
  sample_n(10000) %>%
  ggplot(aes(x = age, y = vote_prop, col = factor(PARTY_CODE))) +
  geom_point(alpha = .3) +
  stat_smooth() +
  xlim(c(0, 100))
```


##

```{r echo = FALSE, warning=FALSE}
voter_sample <- read_csv("http://bit.ly/2oOdqlV")
voter_tidy <- read_csv("http://bit.ly/2I9CFaC")
library(lubridate)
voter_tidy %>%
  mutate(birth_year = year(mdy(BIRTH_DATE)),
         age = 2016 - birth_year) %>%
  filter(PARTY_CODE %in% c("REP", "DEM", "NAV")) %>%
  sample_n(10000) %>%
  ggplot(aes(x = age, y = vote_prop, col = factor(PARTY_CODE))) +
  geom_point(alpha = .3) +
  stat_smooth() +
  xlim(c(0, 100))
```

Is this age?


# Regression

## Focus on the precinct {.build}

Let's go back pre-voter_sample and create a data set where

- unit of observation is the **precinct**
- we're only looking at the 2016 election.

Why would we do this?

- ease of computing
- merge with precinct-level information
- now we're talking about voter turnout


## Focus on the precinct I {.build}

```{r echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE}
library(DescTools)
wd <- getwd()
setwd("../../omv/or_voter_history")

filenames <- list.files(pattern = "2017.txt")
files <- lapply(filenames,read_delim, 
                col_types = paste(rep("c", 65), collapse = ""),
                delim = "\t", 
                escape_double = FALSE, 
                trim_ws = TRUE)

vs <- bind_rows(files)
rm(files)
```


## Focus on the precinct II {.build}

```{r echo = TRUE, eval = FALSE}
or_precinct <- vs %>%
  select(PARTY_CODE, COUNTY, BIRTH_DATE, `11/08/2016`, PRECINCT) %>%
  mutate(age_at_election = (mdy(BIRTH_DATE) %--% 
                            mdy("11/08/2016")) %/% 
                            years(1)) %>%
  group_by(COUNTY, PRECINCT) %>%
  summarize(p_vote    = mean(`11/08/2016` == "YES", na.rm = TRUE),
            avg_age   = mean(age_at_election, na.rm = TRUE),
            p_NAV     = mean(PARTY_CODE == "NAV", na.rm = TRUE),
            party_div = Gini(as.factor(PARTY_CODE)),
            count     = n()) %>%
  mutate(metro = COUNTY %in% c("MULTNOMAH", "CLACKAMAS", 
                               "WASHINGTON", "LANE", "BEND"))
setwd(wd)
write_csv(or_precinct, "or_precinct.csv")
```


##  {.build}

```{r message = FALSE}
or_precinct <- read_csv("../../omv/or_precinct.csv")
head(or_precinct)
```

```{r}
or_precinct <- filter(or_precinct, COUNTY != "ACP")
```

## {.build}

What is the relationship between `p_vote` and `avg_age`?

```{r}
ggplot(or_precinct, aes(x = avg_age, y = p_vote)) +
  geom_point(alpha = .3)
```

## {.build}

Some sensible filtering.

```{r}
or_precinct <- filter(or_precinct, count > 10)
```


## {.build}

What is the relationship between `p_vote` and `avg_age`?

```{r}
ggplot(or_precinct, aes(x = avg_age, y = p_vote)) +
  geom_point(alpha = .3)
```


## {.build .smaller}

What is the relationship between `p_vote` and `avg_age`?

```{r}
m1 <- lm(p_vote ~ avg_age, data = or_precinct)
summary(m1)
```

## {.build}

What is the relationship between `p_vote` and `p_NAV`?

```{r}
ggplot(or_precinct, aes(x = p_NAV, y = p_vote)) +
  geom_point(alpha = .3)
```

## {.build .smaller}

What is the relationship between `p_vote` and `p_NAV`?

```{r}
m2 <- lm(p_vote ~ p_NAV, data = or_precinct)
summary(m2)
```

## {.build}

What is the relationship between `p_vote` and `metro`?

```{r}
ggplot(or_precinct, aes(x = metro, y = p_vote)) +
  geom_boxplot()
```

## {.build .smaller}

What is the relationship between `p_vote` and `metro`?

```{r}
m3 <- lm(p_vote ~ metro, data = or_precinct)
summary(m3)
```

## {.build}

What is the relationship between `p_vote` and `party_div`?

```{r warning = FALSE}
ggplot(or_precinct, aes(x = party_div, y = p_vote)) +
  geom_point(alpha = .3)
```

## {.build .smaller}

What is the relationship between `p_vote` and `party_div`?

```{r}
m4 <- lm(p_vote ~ party_div, data = or_precinct)
summary(m4)
```

## {.build}

What is the relationship between `p_vote` and `p_NAV` when controlling for `metro`?

```{r}
ggplot(or_precinct, aes(x = avg_age, y = p_vote, color = metro)) +
  geom_point(alpha = .3)
```

## {.build .smaller}

What is the relationship between `p_vote` and `p_NAV` when controlling for `metro`?

```{r}
m5 <- lm(p_vote ~ p_NAV + metro, data = or_precinct)
summary(m5)
```

## {.build}

What is the relationship between `p_vote` and `p_NAV` when controlling for `avg_age`?

```{r message = FALSE}
library(plotly)
(p <- plot_ly(or_precinct, x = ~p_NAV, y = ~avg_age, 
        z = ~p_vote, colors = '#BF382A', opacity = .3, marker = list(size = 3)) %>%
  add_markers())
```

## {.build .smaller}

What is the relationship between `p_vote` and `p_NAV` when controlling for `avg_age`?

```{r}
m6 <- lm(p_vote ~ p_NAV + avg_age, data = or_precinct)
summary(m6)
```


## {.build}

```{r message = FALSE, warning = FALSE, echo = FALSE}
library(reshape2)
#Graph Resolution (more important for more complex shapes)
graph_reso <- 0.05

#Setup Axis
axis_x <- seq(0, 1, by = graph_reso)
axis_y <- seq(min(or_precinct$avg_age), max(or_precinct$avg_age), by = graph_reso)

#Sample points
lm_surface <- expand.grid(p_NAV = axis_x, avg_age = axis_y, KEEP.OUT.ATTRS = F)
lm_surface$p_vote <- predict.lm(m6, newdata = lm_surface)
lm_surface <- acast(lm_surface, avg_age ~ p_NAV, value.var = "p_vote")

add_trace(p,
          z = lm_surface,
          x = axis_x,
          y = axis_y,
          type = "surface") %>% layout(showlegend = FALSE)
```


# Classification


## Example: Credit Default

```{r}
library(ISLR)
head(Default)
```


## Exploratory Data Analysis

```{r, echo = FALSE, fig.height = 3.5}
library(ggplot2)
library(gridExtra)
p1 <- ggplot(Default, aes(x = default, y = balance)) +
  geom_boxplot() + 
  coord_flip() +
  theme_bw()

p2 <- ggplot(Default, aes(x = default, y = income)) +
  geom_boxplot() + 
  coord_flip() +
  theme_bw()

grid.arrange(p1, p2, ncol=2)
```


## Model Fitting

```{r}
m1 <- glm(default ~ balance, data = Default, family = binomial)
coef(m1)
```


## Logistic Model

$$ \hat{P}(Y = 1) = \frac{1}{1 + e^{-(-10.65 + 0.0055 x_i)}} $$

```{r, echo = FALSE, message=FALSE}
library(dplyr)
Default <- mutate(Default, defaultYes = ifelse(default == "Yes", 1, 0))
  
p3 <- ggplot(Default, aes(x = balance, y = defaultYes)) +
  geom_point(pch = 1, alpha = .3, color = "steelblue") + 
  ylab("probability of default") + 
  theme_bw()

p3
```


## Logistic Model

$$ \hat{P}(Y = 1) = \frac{1}{1 + e^{-(-10.65 + 0.0055 x_i)}} $$

```{r, echo = FALSE}
balance <- seq(0, max(Default$balance), length.out = 500)
y <- predict(m1, newdata = data.frame(balance), type = "response")
p_y <- data.frame(balance, y)

p3 + geom_line(data = p_y, aes(x = balance, y = y), lwd = 1.3, col = "grey45")
```

## 
```{r}
summary(m1)$coef
```

Where did those SEs come from?


# Multiple Logisitic Regression

## Add student as a predictor?

```{r echo=FALSE}
ggplot(Default, aes(x = student, y = balance)) +
  geom_boxplot() + 
  coord_flip() +
  theme_bw()
```


##

```{r}
m2 <- glm(default ~ balance + student, data = Default, family = binomial)
summary(m2)$coef
```

What's going on?


##

```{r echo = FALSE}
balance <- rep(seq(0, max(Default$balance), length.out = 500), 2)
student <- c(rep("Yes", 500), rep("No", 500))
y <- predict(m2, newdata = data.frame(balance, student), type = "response")
p_y <- data.frame(balance, student, y)

p3 + geom_line(data = p_y, aes(x = balance, y = y, color = student), lwd = 1.3)
```


# Activity

## Oregon voters

Using `http://bit.ly/2oOdqlV` (voter_sample.csv), build a model to answer the following question:

What is the association between age and whether or not someone voted in the November 2016 election, after controlling for political party?






