---
title: "Tidy Data Activity"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

On the following page, there are 4 representations of the same data set concerning occurances of Tuberculosis in various countries.

\begin{enumerate}
\item For each of the representations, what are the \emph{observations} or \emph{cases}? These refer to specific, unique, and similar sorts of things.  What are the \emph{variables}? These should take the same sort of \emph{value} for each case.

\vspace{9.5cm}

\item Let's say you wanted to create a plot to visualize the change in the rate of TB infection in each country from 1999 to 2000. For each representation, how many \texttt{dplyr} commands would be needed to create a \texttt{rate} variable? How many commands would be needed to transform the data so that you could visualize the raw counts of TB cases in each country from 1999 to 2000?

\end{enumerate}

\newpage\null\thispagestyle{empty}\newpage

```{r echo = FALSE}
library(tidyverse)
A <- table3
B <- table2
C <- table1
D1 <- table4a
D2 <- table4b
```

```{r}
A
B
C
```

\newpage

```{r}
D1
D2
```

\newpage

# Tidy Data Practice

For each of the following data sets,

- decide on the type of reshape needed (\texttt{gather()} or \texttt{spread()}),
- record the \emph{key} and \emph{value}, and 
- sketch what the tidied data set would look like.

\vspace{1cm}

\includegraphics[width = 5in]{../figs/tidy-religion}

\newpage

\includegraphics[width = 5in]{../figs/tidy-weather}

\vspace{5.5cm}

\includegraphics[width = 5in]{../figs/tidy-billboard}



